---
title: "Task3"
author: "Diana Shchuka"
format: 
  html:
    embed-resources: true
editor: source
---

1. Add dataset from file airquality.csv and rename to "df".

```{r}
df <- airquality
```

2. What are the column names of the data frame?

```{r}
colnames(df)
```

3. What are the row names of the data frame?

```{r}
rownames(df)
```

4. Extract the first 6 rows of the data frame and print them to the console

```{r}
head(df, 6)
```

5. How many observations (i.e. rows) are in this data frame?

```{r}
nrow(df)
```

6. Extract the last 6 rows of the data frame and print them to the console.

```{r}
tail(df, 6)
```

7. How many missing values are in the “Ozone” column of this data frame?

```{r}
sum(is.na(df$Ozone))
```

8. What is the mean of the “Ozone” column in this dataset? Exclude missing values (coded as NA) from this calculation.

```{r}
mean(df$Ozone, na.rm = TRUE)
```

9. Extract the subset of rows of the data frame where Ozone values are above 31 and Temp values are above 90.

```{r}
subset_data <- df[df$Ozone > 31 & df$Temp > 90, ]
print(subset_data)
```

