---
title: "Task2"
author: "Diana Shchuka"
format: 
  html:
    embed-resources: true
editor: source
---

1. Create the next random vector:

v <- rnorm(50)

For the vector print:

10th item,
items from 10th to 20th (including 20th item),
10 items starting from 5th,
all items bigger than 0,
last 5 items.

```{r}
# Create a random vector of 50 items from a normal distribution
set.seed(123)  # Set seed for reproducibility
v <- rnorm(50)

# Print the 10th item
tenth_item <- v[10]
cat("10th item:", tenth_item, "\n")

# Print items from 10th to 20th (including 20th item)
items_10_to_20 <- v[10:20]
cat("Items from 10th to 20th:", items_10_to_20, "\n")

# Print 10 items starting from the 5th
items_5_to_14 <- v[5:(5+9)]
cat("10 items starting from 5th:", items_5_to_14, "\n")

# Print all items bigger than 0
items_bigger_than_0 <- v[v > 0]
cat("All items bigger than 0:", items_bigger_than_0, "\n")

# Print the last 5 items
last_5_items <- tail(v, 5)
cat("Last 5 items:", last_5_items, "\n")
```

2. Create the next data frame:

df <- 
  data.frame(a = rnorm(50), b = 1:50, 
             cc = sample(letters, 50, replace = TRUE))
             
For the data frame find and print:

the last 10 rows,
15th item of columns cc,
subset columns a and b and rows from 10th to 20th (including 20th),
a new vector from column b.

```{r}
# Create the data frame
set.seed(123)  # Set seed for reproducibility
df <- data.frame(
  a = rnorm(50),                        # Column a with random normal values
  b = 1:50,                             # Column b with sequential numbers from 1 to 50
  cc = sample(letters, 50, replace = TRUE)  # Column cc with random letters
)

# Print the last 10 rows
cat("Last 10 rows:\n")
print(tail(df, 10))

# Print the 15th item of column cc
fifteenth_item_cc <- df$cc[15]
cat("15th item of column cc:", fifteenth_item_cc, "\n")

# Print subset of columns a and b and rows from 10th to 20th (including 20th)
cat("Subset of columns a and b (rows 10 to 20):\n")
subset_ab <- df[10:20, c("a", "b")]
print(subset_ab)

# Create a new vector from column b
new_vector_b <- df$b
cat("New vector from column b:\n")
print(new_vector_b)
```

3. Create the next random vector:

v_na <- rnorm(30)
num_na <- sample(10, 1)
v_na[sample(30, num_na)] <- NA

For this vector:

print all not NA items,
find how many NA values the vector has,
find min max and mean for the vector.

```{r}
# Create a random vector with 30 items from a normal distribution
set.seed(123)  # Set seed for reproducibility
v_na <- rnorm(30)

# Randomly select a number of NA values to insert
num_na <- sample(10, 1)
v_na[sample(30, num_na)] <- NA  # Insert NA values into the vector

# Print all non-NA items
cat("All non-NA items:\n")
print(v_na[!is.na(v_na)])

# Find how many NA values the vector has
na_count <- sum(is.na(v_na))
cat("Number of NA values:", na_count, "\n")

# Find min, max, and mean for the vector (excluding NA values)
min_value <- min(v_na, na.rm = TRUE)
max_value <- max(v_na, na.rm = TRUE)
mean_value <- mean(v_na, na.rm = TRUE)

cat("Minimum value (excluding NA):", min_value, "\n")
cat("Maximum value (excluding NA):", max_value, "\n")
cat("Mean value (excluding NA):", mean_value, "\n")
```